---
title: "FormattableTables"
author: "Elliott Schmidt"
date: "11/03/2022"
output:
  html_document: 
    theme: spacelab
    toc: true
    toc_float: true  
    collapsed: true
    smooth_scroll: true
fontsize: 12pt
mainfont: Arial
monofont: DejaVu Sans Mono
classoption: a4paper
---

_"And why should we feel anger towards the world?_
_As if the world would notice"_

_-Marcus Aurelius_

# Setup 

Download necessary packages  

```{r packages_needed, message=FALSE, warning=FALSE} 
library(usethis) 
library(dataaimsr) 
library(purrr)
library(tidyverse)
library(lubridate)
library(reshape2)
library(formattable)
library(DT)
library(ggplot2)
library(kableExtra)
```
Before beginning you need to make sure that your **AIMS API Key** is hidden in your a **.Renviron** file. To create an .Renviron file you can use the following code. 

```{r renviron_file, eval=FALSE}
usethis::edit_r_environ()
```
Within your newly create .Renviron file enter your AIMS API Key as AIMS_DATAPLATFORM_API_KEY=XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX 

This will prevent you from accidentally pushing your AIMS API Key onto GitHub. Your AIMS API Key is meant just for you, donot put it on GitHub. To check that your AIMS API Key is stored correctly run 

```{r aims_api_key1, results="hide"}
Sys.getenv("AIMS_DATAPLATFORM_API_KEY")
```
If your AIMS API Key shows up, then it is stored correctly. But we don't want to be typing it out every time. Therefore hide it in a new variable called **my_api_key**

```{r aims_api_key2}
my_api_key <- Sys.getenv("AIMS_DATAPLATFORM_API_KEY") 
```

# Importing data from AIMS  

Lets looks at what attributes we can filter by
```{r importing_data}
dataaimsr::aims_expose_attributes("temp_loggers") 
```

Lets look at a summary of data by reef
```{r importing_data2, message=FALSE, results='hide'}
summary_series_data <- aims_data("temp_loggers", api_key = my_api_key,
          summary = "summary-by-series")
head(summary_series_data)
```
```{r importing_data3, echo=FALSE}
kable(as.data.frame(head(summary_series_data)), "html") %>% 
  kable_styling("striped") %>% 
  scroll_box(width = "100%")
```


Lets try to get a summary of daily temperatures from Davies Reef and set an interval of dates 
```{r importing_data4, message=FALSE, results='hide'}
davies <- aims_data("temp_loggers", 
                       api_key = my_api_key, 
                       summary = "daily",
                       filters = list(
                         "series" = "DAVSL1",
                         "from_date" = "2010-03-16T00:00:00",
                         "thru_date" = "2020-01-07T00:00:00"
)) ; head(davies)
```

```{r importing_data5, echo=FALSE}
kable(head(davies), "html") %>% 
  kable_styling("striped") %>% 
  scroll_box(width = "100%")
```

# Selecting data by region 
```{r selection_regional_data, message=FALSE}
CairnsRegion <- aims_data("temp_loggers", 
                       api_key = my_api_key,
                       summary = "summary-by-series", 
                       filters = list(max_lat = -0.000, 
                                      min_lat = - 12.000,
                                      max_lon = 150.500, 
                                      min_lon = 143.500, 
                                      from_date = "2016-01-01", 
                                      thru_date = "2020-12-31"
                                      )) ; CairnsRegion
```
## Output {.tabset} 

### List 
```{r output_list, echo=FALSE}
unique(CairnsRegion$site)
```

### Map
```{r output_map, echo=FALSE}
plot(CairnsRegion, ptype = "map") 
```

### Map (corrected)
```{r output_map2, echo=FALSE, message=FALSE}
CairnsRegion2 <- aims_data("temp_loggers", 
                          api_key = my_api_key,
                          summary = "summary-by-series", 
                          filters = list(max_lat = -15.000, 
                                         min_lat = - 19.000,
                                         max_lon = 147.500, 
                                         min_lon = 145.500, 
                                         from_date = "2015-01-01", 
                                         thru_date = "2020-12-31"
                          ))  ; plot(CairnsRegion2, ptype = "map") 
```


# Data Manipulation
```{r data_manipulation}
#--- filtering data ---#
CairnsRegion3 <- CairnsRegion2%>% 
  filter(between (depth, 7,15)) 

#--- summarising data ---#
CairnsRegion3 %>% summarise( 
  number_of_reefs = length(site), 
  min_depth = min(depth), 
  max_depth = max(depth), 
  mean_depth = mean(depth)
  )

unique(CairnsRegion3$site)
```

The default for max. observations downloaded per site is set to 1000. If downloading a large amount of data set the **size** parameter to a large number 

Lets start by downloading data from **2015-01-01** to **2020-12-31** for the reefs that we identified as being in Cairns Region earlier. 

The list of reefs will be store in variable **reef_list**
```{r data_manipulation2, message=FALSE, cache=TRUE}
#--- Importing data from multiple sites ---#
reef_list <- CairnsRegion3$series_id
aims_data_per_series <- function(series_number, my_api_key, ...) {
  aims_data("temp_loggers", api_key = my_api_key,  summary = "daily",
            filters = list(series_id = series_number, ...))
}

CairnsTemp <- purrr::map_df(reef_list, aims_data_per_series,
                         my_api_key = my_api_key,
                         from_date = "2015-01-01",
                         thru_date = "2020-12-31", 
                         size = 3000) 
```

Lets look at a quick summary of out data 

```{r data_manipulation3, message=FALSE} 
CairnsTemp_summary <- CairnsTemp %>% 
  group_by(site) %>% 
  summarise( 
    earliest_date = min(time), 
    latest_date = max(time), 
    days_obsrvd = length(time))
print(as_tibble(CairnsTemp_summary), n=length(CairnsTemp_summary$site)) 
```

Great Job! But to make our lives easier before we move on, will we separate the time column into the variables **YEAR**,**MONTH**, and **DAY**. 

```{r data_manipulation4}
CairnsTemp2 <- CairnsTemp %>% 
  mutate(time = as_date(time)) %>% #reformat date column as a 'date' variable 
  separate(time, 
           sep="-", 
           remove=FALSE, 
           into = c("YEAR", "MONTH", "DAY")) %>%  
  mutate(YEAR = as.numeric(YEAR), 
         MONTH = as.numeric(MONTH), 
         DAY = as.numeric(DAY))

# quick check to see if it worked 

names(CairnsTemp2)
``` 

Lets save this dataframe. Although it would be easy enough for potential collaborators to run this script image you were working with a large dataframe and/or model that was slow to work with. It will save you time by not having to download all the AIMS temperature logger data everytime you want to work on it as well.  

Saving dataframes is easy. We will save the dataframe in a new directory called **'files'**. To do so, uncomment the first line of code below.  

```{r saving_data_frame}
#dir.create("files")
save(CairnsTemp2, file="./files/CairnsTemp2.Rda")
```

# Formating dataframe for FormatTable 

Below is how you would upload your saved dataframe
```{r uploading_saved_dataframe}
#load dataframe
load("./files/CairnsTemp2.Rda")  
``` 

```{r formating_data_formattable, message=FALSE}
#--- formatting data for table ---# 
#extracting summer months
summer <- CairnsTemp2 %>%
  filter(MONTH == c(1,2,3,12)); unique(summer$MONTH) 

#If you want to save, uncomment lines below
#save(summer, file = "./files/summer.Rda") 

#--- gather yearly means ---#
mst <- summer %>% 
  group_by(site, YEAR)%>% 
  na.omit() %>%
  summarise(temp_mean = mean(cal_val), 
             max_mean = mean(cal_max), 
             depth_mean = mean(depth), 
             range_mean = mean(cal_max - cal_min)) %>% 
  ungroup() 
head(mst)
```

The function **dcast** does a lot of heavy lifting of us, so we must be kind to it. **dcast** will reformat the data to by listing 'sites' within rows and 'Year' as the columns. Lets run the code and see what it looks like.   

We will collect data for temperature **mean**, **max**, **range**, and logger **depth** separately and then bring all the data back together. 

```{r formating_data_formattable2}
#--- getting data into table format in R ---# 
mst_temp <- mst %>% 
  dcast(site~YEAR, value.var = "temp_mean") 
colnames(mst_temp)[2:length(colnames(mst_temp))] = paste0('Mean_',colnames(mst_temp)[2:length(colnames(mst_temp))])

mst_max <- mst %>% 
  dcast(site~YEAR, value.var = "max_mean")
colnames(mst_max)[2:length(colnames(mst_temp))] = paste0('Max_',colnames(mst_max)[2:length(colnames(mst_temp))])

mst_depth <- mst %>% 
  dcast(site~YEAR, value.var = "depth_mean")
colnames(mst_depth)[2:length(colnames(mst_temp))] = paste0('Depth_',colnames(mst_depth)[2:length(colnames(mst_temp))])

mst_range <- mst %>% 
  dcast(site~YEAR, value.var = "range_mean") 
colnames(mst_range)[2:length(colnames(mst_temp))] = paste0('Range_',colnames(mst_range)[2:length(colnames(mst_temp))])

head(mst_temp) 

# join all the data together
mst_all <- full_join(mst_temp,mst_max, by="site") %>% 
  full_join(mst_range, by ="site") %>% 
  full_join(mst_depth[c(1,7)], by = "site") %>% 
  mutate_if(is.numeric, round, digits = 2); names(mst_all) 
#head(mst_all)
#save data
#save(mst_all, file="./files/mst_all.Rda")
```

# FormatTable 

```{r FormatTable}
#--- put table into formattable ---# 
# load data
#load("./files/mst_all.RDA")
mst_all2 <- mst_all
mst_all2[is.na(mst_all2)] = '' 

CairnsRegion_Table <- formattable(mst_all2, 
            align = c("l","r","r","r","r","r","r","r","r","r","r","r","r"
                      ,"r","r","r","r","r","r","r","r","r","r","r","r","r"),
            list(site = formatter( 
              "span",
              style = ~style(color = "grey", font.weigh = "bold")), 
              area(col = `Mean_2015`:`Mean_2020`) ~ color_tile("white","red"), 
              area(col = `Max_2015`:`Max_2020`) ~ color_tile("yellow","red"), 
              area(col = `Range_2015`:`Range_2020`) ~ color_tile("palegreen","green3"),
              Depth_2020 = color_bar("pink", 'proportion', 0.2))) 

CairnsRegion_Table = as.datatable(formattable(CairnsRegion_Table)) %>% 
  formatStyle(colnames(mst_all2), `text-align` = 'right')
#CairnsRegion_Table
```

Save table using the 'export' button in RStudio 

# Density Plots 

## Descriptive statistics 
```{r descriptive statistics}
#--- descriptive statistics ---# 
CairnsTemp4 <- CairnsTemp2 %>% 
  mutate(heatwave_years = case_when( 
    YEAR == "2016" ~ "heatwave_yr", 
    YEAR == "2017" ~ "heatwave_yr", 
    YEAR == "2020" ~ "heatwave_yr",
    TRUE ~ "ambient_yr"))

agg_data <- CairnsTemp4 %>% 
  group_by(heatwave_years)%>% 
  na.omit() %>%
  summarise(temp_mean = mean(cal_val), 
            max_mean = mean(cal_max), 
            range_mean = mean(cal_max - cal_min)) %>% 
  ungroup() 

agg_data[nrow(agg_data)+1,] <- list("Difference",
                                    agg_data$temp_mean[2]-agg_data$temp_mean[1], 
                                    agg_data$max_mean[2]- agg_data$max_mean[1], 
                                    agg_data$range_mean[2] - agg_data$range_mean[1])
agg_data  

# But his doesn't help us with determining summer temperature 

#load("./files/summer.Rda")

CairnsTemp5 <- summer %>% 
  mutate(heatwave_years = case_when( 
    YEAR == "2016" ~ "heatwave_yr", 
    YEAR == "2017" ~ "heatwave_yr", 
    YEAR == "2020" ~ "heatwave_yr",
    TRUE ~ "ambient_yr"))

agg_data <- CairnsTemp5 %>% 
  group_by(heatwave_years)%>% 
  na.omit() %>%
  summarise(temp_mean = mean(cal_val), 
            max_mean = max(cal_max), 
            range_mean = mean(cal_max - cal_min)) %>% 
  ungroup() 

agg_data[nrow(agg_data)+1,] <- list("Difference",
                                    agg_data$temp_mean[2]-agg_data$temp_mean[1], 
                                    agg_data$max_mean[2]- agg_data$max_mean[1], 
                                    agg_data$range_mean[2] - agg_data$range_mean[1])
agg_data 
```

## Density plot 

**Note: Make sure your colors are friendly to the colorblind!** 

```{r plotting} 
my_temperature_plot <- CairnsTemp4 %>% 
  na.omit() %>% 
  ggplot(aes(cal_val))+
  geom_density(aes(fill = heatwave_years), 
               size = 1, 
               position = "identity", 
               alpha = 0.8, 
               adjust = 1.5) + 
  scale_fill_manual(labels = c("Ambient", "Heatwave"), 
                    values = c("#56B4E9", "#D55E00")) + 
  scale_x_continuous(breaks = seq(21,34,2), 
                     limits = c(21,32))+
  theme_light()+ 
  xlab("Temperature (\u00B0C)") + 
  ylab("Density")+ 
  ylim(0,0.25)+
  ggtitle("My temperature plot :)")+ 
  labs(fill = "Temperature years") + 
  theme(legend.direction = "horizontal",
        legend.position = c(.25,.95)) + 
  guides(shape = guide_legend(override.aes = list(size = 0.2))); my_temperature_plot

# what are the exact temperature with the max. frequency

ggplot_build(my_temperature_plot)$data[[1]] %>%
  group_by(fill) %>% 
  filter(density == max(density)) %>% 
  ungroup() %>% 
  kable("html") %>% 
  kable_styling("striped") %>% 
  scroll_box(width = "100%")

``` 

Save the plot. There are one or two way of doing this 

```{r saving plot, eval=FALSE}
###saving plot####
ggsave(filename = "./files/my_temperature_plot.pdf", 
       width = 8, 
       height = 6, 
       device = 'pdf', 
       dpi = 500)
dev.off()

#OR 

pdf(file = "./files/my_temperature_plot2.pdf", 
    width = 8, 
    height = 6)
my_temperature_plot 
dev.off() 
```

You can also get the same information from the dataframe 

```{r data extraction, eval=TRUE}
ambient_temp_summer <- CairnsTemp4 %>% 
  filter(heatwave_years == "ambient_yr") %>% 
  mutate_if(is.numeric, round, digits = 2) %>%
  {.$cal_val} %>% 
  table() %>% 
  which.max() %>% 
  names() 
heatwave_temp_summer <- CairnsTemp4 %>% 
  filter(heatwave_years == "heatwave_yr") %>% 
  mutate_if(is.numeric, round, digits = 2) %>%
  {.$cal_val} %>% 
  table() %>% 
  which.max() %>% 
  names()
``` 
```{r data exatraction2, echo = FALSE}
df <- data.frame(ambient_temp_summer = ambient_temp_summer, 
                 heatwave_temp_summer = heatwave_temp_summer); df
```

Why are the numbers from the plot and dataframe different?

```{r data_by_10, eval=FALSE}
#--- 10 minute data instead of daily ---#
davies <- aims_data("temp_loggers", 
                    api_key = my_api_key,
                    filters = list(
                      series = "DAVSL1",
                      from_date = "2013-01-14",
                      thru_date = "2013-01-28"
                    )) ; head(davies)
```